{How do we copy and paste?\\The semantic drift of quotations in blogspace}

### 2014-12-12 brains-introduction.tex
US spelling, and phrasing, for first section (introduction)

The understanding of knowledge transmission mechanisms has led to a sizeable literature in the recent past, spanning over numerous research fields ranging from cultural anthropology to social network analysis and complex systems modeling, from social cognition to data mining. These works are diversely labeled as studies on ``opinion dynamics'', ``cultural evolution'', or ``information diffusion'' and, for the most part, 
%Broadly speaking, this type of research program is bound to
 investigating phenomena pertaining to both cognitive science and social science, %and aims to understand the processing, transmission, and evolution of information 
 both at the individual and social levels.  Broadly speaking, we may distinguish two main research streams, depending on whether the focus lies on cognitive processes or on social dynamics.

@@

A first stream of research is largely structured around cultural anthropology and essentially addresses cultural similarity, diversity and its evolution. It features several theories mixing social and individual cognition including, to cite a few, the debate around the ``memetic'' program initiated by \citet{Dawkins76} (for which the collection of works by \citet{Aunger00} provides a solid overview) the development of evolutionary models of norms (see for instance \citet{Ehrlich05}, following the seminal work of \citet{Boyd85}); or the program of ``cultural epidemiology'', initiated by \citet{sper:expl}, which %recently attracted significant attention, and
% focuses on the notion of \emph{representation}, 
links %the cognitive science 
concept of \emph{mental representation} to the concept of \emph{public representation} (the latter being the former's counterpart outside the brain, i.e. in cultural artifacts: texts, utterances, etc.).\footnote{\citet{sper:expl} emphasizes this distinction in his seminal work:
\begin{quotation}
A representation may exist inside its user: it is then a \emph{mental representation}, such as a memory, a belief, or an intention.
The producer and the user of a mental representation are one and the same person.
A representation may also exist in the environment of its user, as is the case, for instance, of the text you are presently reading: it is then a \emph{public representation}.
\end{quotation}}

Here, one of the core claim consists in observing that not all knowledge is equally fit for being reproduced, although the various approaches have a different take on how exactly this notion of fitness should be operationalized. Sperber's cultural epidemiology opposes Dawkins' memetics by insisting that representations are not being replicated through a high-fidelity copy process, but are being interpreted and produced anew, and are thus greatly subject to change.
Cultural epidemiology postulates this type of conceptual evolution and suggests that it can be appraised through the notion of ``cultural attractor'', seen as the attraction domain of an underlying socio-semantic dynamical system.\footnote{Works such as \citet{Atran03} argue that this approach is anthropologically better suited than memetics, and the main issues in this debate are further detailed by \citet{Kuper00} and \citet{Bloch00}.}
Despite some recent modelling attempts (for instance \citet{Claidiere07}), the development of quantitative measurements focused on the key notion of cultural attractor has remained a relatively hard task and, to our knowledge, this hypothesis has not yet been extensively empirically analyzed.

Another research stream deals with rather macroscopic studies of knowledge diffusion. Here, one of the focal points is that not all knowledge gets propagated identically along the same routes, within the same communities, at the same pace. The various approaches usually feature a minimalistic description of cognitive processes, strongly reminiscent of biological epidemiology (a single, atomic piece of information may or may not be adopted by individuals).  
This research program nonetheless exhibits a significant empirical track record --- largely owing to a recent avalanche of observable \emph{in vivo} data which, for a good decade now, has mainly come from a variety of online interaction contexts.
While these information trails are not records of ``physical'' inter-individual interactions (in the sense of ``real life'' interactions), they still constitute a wealth of observations on the dynamics of public --~albeit online~-- representations.
Some could describe for instance the propagation of cultural artifacts across social networks such as blogspace \citep{Gruhl04}, the characteristic times and diffusion cycles both within these social networks and with respect to the topical dynamics of news media \citep{Leskovec09}, or the reciprocal influence between the social network topology and the distribution of issues \citep{Cointet09}.

These studies are relatively remote from anthropology and cognition, and, for a significant part, rely only marginally on specific social science theories; they are rather are at the interface between data mining, complex systems and quantitative sociology (first and foremost social network analysis). They nonetheless show us the interest of using these records --~given the already significant and rapidly growing importance of our online interactions~--
%and, %more broadly, large textual corpora \cam{google n-grams??} 
to dramatically improve the empirical understanding of individual-level processes of cultural evolution.

>>>>

The understanding of knowledge transmission mechanisms has led to a sizable literature in the recent past, spanning over numerous research fields ranging from cultural anthropology to social network analysis and complex systems modeling, from social cognition to data mining. These works are diversely labeled as studies on ``opinion dynamics'', ``cultural evolution'', or ``information diffusion'' and, for the most part, 
%Broadly speaking, this type of research program is bound to
 investigate phenomena pertaining to both cognitive science and social science, %and aims to understand the processing, transmission, and evolution of information 
 both at the individual and social levels.  Broadly speaking, we may distinguish two main research streams, depending on whether the focus lies on cognitive processes or on social dynamics.

@@

A first stream of research is largely structured around cultural anthropology and essentially addresses cultural similarity, diversity and its evolution. It features several theories mixing social and individual cognition including, to cite a few, the debate around the ``memetic'' program initiated by \citet{Dawkins76} (for which the collection of works by \citet{Aunger00} provides a solid overview), the development of evolutionary models of norms (see for instance \citet{Ehrlich05}, following the seminal work of \citet{Boyd85}); or the ``cultural epidemiology'' program initiated by \citet{sper:expl}, which %recently attracted significant attention, and
% focuses on the notion of \emph{representation}, 
links %the cognitive science 
the concept of \emph{mental representation} to the concept of \emph{public representation} (the latter being the former's counterpart outside the brain, i.e. in cultural artifacts: texts, utterances, etc.).\footnote{\citet{sper:expl} emphasizes this distinction in his seminal work:
\begin{quotation}
A representation may exist inside its user: it is then a \emph{mental representation}, such as a memory, a belief, or an intention.
The producer and the user of a mental representation are one and the same person.
A representation may also exist in the environment of its user, as is the case, for instance, of the text you are presently reading: it is then a \emph{public representation}.
\end{quotation}}

Here, one of the core claims consists in observing that not all knowledge is equally fit for being reproduced, although the various approaches have a different take on how exactly this notion of fitness should be operationalized. Sperber's cultural epidemiology opposes Dawkins' memetics by insisting that representations are not being replicated through a high-fidelity copy process, but are being interpreted and produced anew, and are thus greatly subject to change.
Cultural epidemiology postulates this type of conceptual evolution and suggests that it can be appraised through the notion of ``cultural attractor'', seen as the attraction domain of an underlying socio-semantic dynamical system.\footnote{Works such as \citet{Atran03} argue that this approach is anthropologically better suited than memetics, and the main issues in this debate are further detailed by \citet{Kuper00} and \citet{Bloch00}.}
Despite some recent modeling attempts (for instance \citet{Claidiere07}), the development of quantitative measurements focused on the key notion of cultural attractor has remained a relatively hard task and, to our knowledge, this hypothesis has not yet been \TB{extensively} empirically analyzed.\marginnote{\seb{Deux adverbes. Pas encore trouvé d'autre formulation.}}

Another research stream deals with rather macroscopic studies of knowledge diffusion. Here, one of the focal points is that not all knowledge gets propagated identically along the same routes, within the same communities, at the same pace. The various approaches usually feature a minimalistic description of cognitive processes, strongly reminiscent of biological epidemiology (a single, atomic piece of information may or may not be adopted by each individual).  
This research program nonetheless exhibits a significant empirical track record --- largely owing to a recent avalanche of observable \emph{in vivo} data which, for a good decade now, have mainly come from online interaction contexts.
While these information trails are not records of ``physical'' inter-individual interactions (in the sense of ``real life'' interactions), they still constitute a wealth of observations on the dynamics of public --~albeit online~-- representations.
Some could describe for instance the propagation of cultural artifacts across social networks such as blogspace \citep{Gruhl04}, the characteristic times and diffusion cycles both within these social networks and with respect to the topical dynamics of news media \citep{Leskovec09}, or the reciprocal influence between the social network topology and the distribution of issues \citep{Cointet09}.

These studies are relatively remote from anthropology and cognition, and, for a significant part, rely only marginally on specific social science theories; they are rather at the interface between data mining, complex systems and quantitative sociology (first and foremost social network analysis). They nonetheless show us the interest of using these records --~given the already significant and rapidly growing importance of our online interactions~--
%and, %more broadly, large textual corpora \cam{google n-grams??} 
to dramatically improve the empirical understanding of individual-level processes of cultural evolution.
